
```{r}
data_folder <- here::here("Data", "National_Data")
code_folder <- here::here("Code", "Analysis")
figure_folder <- here::here("Figures")
source(here::here(code_folder, "helper_functions.R"))

library(epidatr)
# library(stats)
library(patchwork)
library(ggplot2)
```


## Load finalized data


```{r}
# Load finalized hospitalizations and deaths, both pre-smoothed
max_jhu_date <- as.Date("2023-03-01") # Not reliable after this
jhu_deaths <- readRDS( here::here(data_folder, "JHU_finalized.RData"))
death_dates <- as.Date(names(jhu_deaths))

first_hosp_date <- as.Date("2020-07-28")
hhs_hosps <- readRDS( here::here(data_folder, "HHS_finalized.RData"))
hosp_dates <- as.Date(names(hhs_hosps))

# Load real-time hospitalizations and deaths
hhs_hosps_realtime <- readRDS(here::here(data_folder, "HHS_real_time.RData"))
jhu_deaths_realtime <- readRDS(here::here(data_folder, "JHU_real_time.RData"))


# Load NHCS HFRs
nhcs_hfrs_all <- readRDS(here::here(data_folder, "HFRs_NHCS_rescaled.RData"))
nhcs_hfr_dates <- as.Date(names(nhcs_hfrs_all))

```



## Get hyperparameters for estimating HFRs

```{r}
# Lag parameter for lagged method
oracle_lag <- compute_optimal_lag(hhs_hosps, jhu_deaths)

# Get delay distribution for Nishiura method
# UK delay distribution
Ns <- c(28611, 1700, 16641, 16903)
means <- c(10.20, 14.06, 14.84, 9.24)
sds <- c(9.51, 14.31, 12.35, 7.64)
gamma_mean_uk <- sum(means*Ns/sum(Ns)) # Weighted average hosp-to-death time
gamma_sd_uk <- sum(sds*Ns/sum(Ns)) # Weighted average of SDs
round(c(gamma_mean_uk, gamma_sd_uk, gamma_sd_uk/gamma_mean_uk), 2)

d <- 75
gamma_mean <- oracle_lag
gamma_sd <- round(gamma_mean*.9)
delay_distr <- make_delay_distr(gamma_mean, gamma_sd, d)

```

## Estimate true HFRs, and smooth approx. ground truth with finalized counts

```{r}
est_weeks <- seq(first_hosp_date+d+14, min(max(hosp_dates)-4*7, max(nhcs_hfr_dates)), by=7)
n_ests <- length(est_weeks)

# Smooth GT HFRs and align with est_weeks
hfrsNHCS <- sapply(est_weeks, function(date) {
  mean(nhcs_hfrs_all[nhcs_hfr_dates %in% seq(date - 6, date, by="day")], na.rm = TRUE)
}); names(hfrsNHCS) <- est_weeks

hfrsHJl <- compute_lagged_hfrs(hhs_hosps, jhu_deaths, l=oracle_lag, w=7, dates=est_weeks)
hfrsHJn <- compute_conv_hfrs(hhs_hosps, jhu_deaths, est_weeks, delay_distr, w=7)

```


## Repeat on properly-versioned real-time data

Figure 3a

```{r}
# Identify valid weeks. 
n_days_after <- 2 # Most recent date often not posted for over a week
week_st_idx <- 5 # HHS first issued Nov 16, 2020 (est_week[4]+1); 
est_weeks_realtime <- est_weeks[week_st_idx:n_ests] 

hfrsNHCS_rt <- hfrsNHCS[week_st_idx:n_ests]

oracle_lag_realtime <- compute_optimal_lag(hhs_hosps_realtime, jhu_deaths_realtime)
hfrsHJl_rt <- compute_lagged_hfrs(hhs_hosps_realtime, jhu_deaths_realtime, 
                                  l=oracle_lag_realtime, w=7, dates=est_weeks_realtime)

delay_distr_realtime <- make_delay_distr(oracle_lag_realtime, round(oracle_lag_realtime*0.9), d)
hfrsHJn_rt <- compute_conv_hfrs(hhs_hosps_realtime, jhu_deaths_realtime, 
                                    est_weeks_realtime, delay_distr_realtime, w=7)

```



```{r}

## Plot just results with real-time counts
HFRtypes <- c("Approx. GT", "Convolutional ratio", "Lagged ratio")
df_real_time <- data.frame(Date=rep(est_weeks_realtime, length(HFRtypes)),
                 HFR=c(hfrsNHCS_rt, hfrsHJn_rt, hfrsHJl_rt), 
                 Method=factor(rep(HFRtypes, each=length(hfrsNHCS_rt)), levels=HFRtypes))

```

## Plot results with both finalized and real-time data

Figure 5 (appendix)

```{r}
n_ests_rt <- length(est_weeks_realtime)
HFRtypes <- c("Ground Truth", "Convolutional", "Lagged")
types_factor <- factor(c(rep(HFRtypes, each=n_ests_rt), rep(HFRtypes[2:3], each=n_ests_rt)), levels=HFRtypes)
Source <- c("GT HFR", "Finalized", "Real-Time")
source_factor <- factor(c(rep(Source[1], each=n_ests_rt), 
                          rep(Source[2:3], each=n_ests_rt*2)), levels=Source)

df_both <- data.frame(Date=rep(est_weeks_realtime, 5),
                 HFR=c(hfrsNHCS_rt, hfrsHJn[week_st_idx:n_ests], 
                       hfrsHJl[week_st_idx:n_ests], hfrsHJn_rt, hfrsHJl_rt),
                 type=types_factor,#rep(as.character(1:5), each=n_ests_rt),
                 source=source_factor
                 )

```




# Robustness checks to choices of hyperparameter

## Robustness to choice of hyperparameter

### Window size


```{r}
# Nishiura convolutional estimator
ws <- c(1, 1:4*7)
est_weeks2 <- est_weeks[4:n_ests]
hfrs_by_w <- lapply(ws, function(w) {
  compute_conv_hfrs(hhs_hosps, jhu_deaths, est_weeks2, delay_distr, w=w)
})

window_labels <- factor(c(rep("Approx. GT", length(est_weeks2)), rep(ws, each=length(est_weeks2))),
                     levels=c("Approx. GT", ws))
df_window_conv <- data.frame(Date=rep(est_weeks2, length(ws)+1),
                 HFR=c(hfrsNHCS[4:n_ests], unlist(hfrs_by_w)),
                 Window=window_labels)
```


```{r}
# Lagged estimator
hfrs_by_w_lag <- lapply(ws, function(w) {
  compute_lagged_hfrs(hhs_hosps, jhu_deaths, l=oracle_lag, w=w, dates=est_weeks2)
})
window_labels <- factor(c(rep("Approx. GT", n_ests-3), rep(ws, each=n_ests-3)),
                     levels=c("Approx. GT", ws))
df_window_lag <- data.frame(Date=rep(est_weeks2, length(ws)+1),
                 HFR=c(hfrsNHCS[4:n_ests], unlist(hfrs_by_w_lag)),
                 Window=window_labels)

```

### Lag

```{r}
# lags <- 2:5*7
lags <- seq(7, 35, by=7)
hfrs_by_lag <- lapply(lags, function(lag) {
  compute_lagged_hfrs(hhs_hosps, jhu_deaths, l=lag, w=7, dates=est_weeks)
})
lag_labels <- factor(c(rep("Approx. GT", n_ests), rep(lags, each=n_ests)),
                     levels=c("Approx. GT", lags))
df_lags <- data.frame(Date=rep(est_weeks, length(lags)+1),
                 HFR=c(hfrsNHCS, unlist(hfrs_by_lag)),
                 Lag=lag_labels)


```


### Choice of delay distribution

```{r}

frac <- 0.9
# frac <- 0.7
gamma_means <- seq(16, 32, by=4)
mat <- matrix(c(gamma_means, round(gamma_means*frac)), nrow=2, byrow = TRUE)
param_list <- as.list(as.data.frame(mat))

delay_shapes <- lapply(param_list, function(mean_and_sd) {
  params <- get_gamma_params(Mean=mean_and_sd[1], Sd=mean_and_sd[2])
  shape <- params[1]; rate <- params[2]
  ddgamma(0:d, shape, rate)})
# delay_shapes <- list(delay_shapes, delay_distr)
hfrs_by_delay <- lapply(delay_shapes, function(delay_shape) {
  compute_conv_hfrs(hhs_hosps, jhu_deaths, est_weeks, delay_shape, w=7)
})

# Find best mean & SD
maes <- sapply(hfrs_by_delay, mae, hfrsNHCS)
round(maes,3)
plot(gamma_means, maes, ylim=c(0, max(maes)))
best_mean <- gamma_means[which.min(maes)]
c(best_mean, round(best_mean*frac))

vector_strings <- unlist(as.character(param_list))
cleaned_numbers <- sapply(vector_strings, function(x) {
  x <- gsub("c\\(|\\)", "", x)
  x <- gsub(", ", "/", x)
})
names(cleaned_numbers) <- NULL
HFRtypes <- c("GT", cleaned_numbers)#as.character(param_list))
df_delays_9 <- data.frame(Date=rep(est_weeks, length(HFRtypes)),
                 HFR=c(hfrsNHCS, unlist(hfrs_by_delay)),
                 Method=factor(rep(HFRtypes, each=n_ests), levels=HFRtypes))
```

```{r}
frac <- 0.7
gamma_means <- seq(16, 32, by=4)
mat <- matrix(c(gamma_means, round(gamma_means*frac)), nrow=2, byrow = TRUE)
param_list <- as.list(as.data.frame(mat))

delay_shapes <- lapply(param_list, function(mean_and_sd) {
  params <- get_gamma_params(Mean=mean_and_sd[1], Sd=mean_and_sd[2])
  shape <- params[1]; rate <- params[2]
  ddgamma(0:d, shape, rate)})
# delay_shapes <- list(delay_shapes, delay_distr)
hfrs_by_delay <- lapply(delay_shapes, function(delay_shape) {
  compute_conv_hfrs(hhs_hosps, jhu_deaths, est_weeks, delay_shape, w=7)
})

# Find best mean & SD
maes <- sapply(hfrs_by_delay, mae, hfrsNHCS)
round(maes,3)
plot(gamma_means, maes, ylim=c(0, max(maes)))
best_mean <- gamma_means[which.min(maes)]
c(best_mean, round(best_mean*frac))

vector_strings <- unlist(as.character(param_list))
cleaned_numbers <- sapply(vector_strings, function(x) {
  x <- gsub("c\\(|\\)", "", x)
  x <- gsub(", ", "/", x)
})
names(cleaned_numbers) <- NULL
HFRtypes <- c("GT", cleaned_numbers)#as.character(param_list))

df_delays_7 <- data.frame(Date=rep(est_weeks, length(HFRtypes)),
                 HFR=c(hfrsNHCS, unlist(hfrs_by_delay)),
                 Method=factor(rep(HFRtypes, each=n_ests), levels=HFRtypes))
```


# Alternative data sources

## Alternative death counts

### Load NCHS deaths and compute HFRs

```{r}
nchs_df <- data.frame(pub_covidcast(source="nchs-mortality",signals= "deaths_covid_incidence_num", 
                                   geo_type = "nation", geo_values="*", 
                                   time_type="week"))

# NCHS reports deaths in week starting at that date
nchs_deaths <- rep(nchs_df$value, each=7)/7
times <- as.Date(nchs_df$time_value)
nchs_dates <- seq(min(times), max(times)+6, by="day")
names(nchs_deaths) <- nchs_dates

# Estimate on dates that signify end of week in NCHS
nchs_week_ends <- times+6
first_est_date <- min(nchs_week_ends[nchs_week_ends > (first_hosp_date + d + 14)])
est_weeks_nchs <- seq(first_est_date, min(max(hosp_dates)-4*7, max(nhcs_hfr_dates), max(nchs_week_ends)), by=7)

```


### Display real-time death counts by source: NCHS, JHU finalized, JHU real-time

```{r}
shared_death_dates <- seq(as.Date("2021-12-15"), as.Date("2022-07-15"), by="day")
jhu_deaths_finalized <- jhu_deaths[names(jhu_deaths) %in% shared_death_dates]
jhu_deaths_realtime2 <- jhu_deaths_realtime[names(jhu_deaths_realtime) %in% shared_death_dates]
nchs_deaths2 <- nchs_deaths[nchs_dates %in% shared_death_dates]
sources <- c("JHU Finalized", "JHU Real time", "NCHS")
df_deaths_by_source <- data.frame(Date=rep(shared_death_dates, 3),
                       Deaths=c(jhu_deaths_finalized, jhu_deaths_realtime2, nchs_deaths2),
                       Source=factor(rep(sources, each=length(shared_death_dates)), levels=sources))
```


## Alternative versions of ground truth HFRs

### Load and compute US variant-based HFRs

```{r}
variant_list <- readRDS( here::here(data_folder, 'variant_hfrs.RData'))
variant_hfrs <- variant_list[['HFRs-2wk lag']]
variant_dates <- variant_list[['Variant Dates']]


n_weeks <- n_ests
const_variant_hfrs <- rep(variant_hfrs[1], n_weeks)
for (i in 1:n_weeks) {
  week <- est_weeks[i]
  if (week >= variant_dates[2] & week < variant_dates[3]) {
    const_variant_hfrs[i] <- variant_hfrs[2] }
  if (week >= variant_dates[3] & week < variant_dates[4]) {
    const_variant_hfrs[i] <- variant_hfrs[3] }
  # if (week >= variant_dates[4]) {
  #   const_variant_hfrs[i] <- variant_hfrs[4] }
  if (week >= variant_dates[4] & week < variant_dates[5]) {
    const_variant_hfrs[i] <- variant_hfrs[5] }
  if (week >= variant_dates[5]) {
    const_variant_hfrs[i] <- variant_hfrs[6] }
}
names(const_variant_hfrs) <- est_weeks_nchs

# Changing HFRs by proportion in circulation
prop_df <- variant_list[['Proportions']]
variant_names <- c("Original", "Alpha", "Delta", "Omicron")
prop_variant_hfrs_all <- as.matrix(prop_df[,variant_names]) %*% variant_hfrs[variant_names]
names(prop_variant_hfrs_all) <- prop_df$Date

prop_variant_hfrs <- approx(as.numeric(prop_df$Date), prop_variant_hfrs_all, as.numeric(est_weeks_nchs))$y
names(prop_variant_hfrs) <- est_weeks_nchs

late_omicron_start <- variant_dates["Late Omicron"]
hfr_start <- as.matrix(prop_df[prop_df$Date < late_omicron_start,variant_names]) %*%
  variant_hfrs[c("Original", "Alpha", "Delta", "Early Omicron")]
hfr_end <- as.matrix(prop_df[prop_df$Date >= late_omicron_start,variant_names]) %*%
  variant_hfrs[c("Original", "Alpha", "Delta", "Late Omicron")]
prop_variant_hfrs_all <- c(hfr_start, hfr_end)
prop_variant_hfrs <- approx(as.numeric(prop_df$Date), prop_variant_hfrs_all, as.numeric(est_weeks_nchs))$y
names(prop_variant_hfrs) <- est_weeks_nchs
```



### Lagged estimates from NHCS deaths


```{r}
hfrsNHCS2 <- sapply(est_weeks_nchs, function(date) {
  mean(nhcs_hfrs_all[nhcs_hfr_dates %in% seq(date - 6, date, by="day")], na.rm = TRUE)
}); names(hfrsNHCS) <- est_weeks_nchs


lag_nchs <- round(gamma_mean_uk); print(lag_nchs)
# lag_nchs <- compute_optimal_lag(hhs_hosps, nchs_deaths, verbose=TRUE)
hfrsHNlretro <- compute_lagged_hfrs(hhs_hosps, nchs_deaths,
                              l=lag_nchs, w=7, dates=est_weeks_nchs, real_time=FALSE, centered=FALSE)

```

### Compare approximate sources of ground truth

```{r}
HFRtypes <- c( "NHCS Survey", "NCHS/HHS (retro)", "GT-Variants")
df_hfr_types <- data.frame(Date=rep(est_weeks_nchs, length(HFRtypes)),
                 HFR=c(hfrsNHCS2, hfrsHNlretro, prop_variant_hfrs), 
                 Method=factor(rep(HFRtypes, each=length(hfrsNHCS)), levels=HFRtypes))

```


## Compare JHU and NCHS real-time Lagged ratios

```{r}
hfrsHNl <- compute_lagged_hfrs(hhs_hosps, nchs_deaths, l=lag_nchs, w=7, dates=est_weeks)
HFRtypes <- c("Approx. GT", "JHU Estimate", "NCHS Estimate")
df_hfr_ests_by_source <- data.frame(Date=rep(est_weeks, length(HFRtypes)),
                 HFR=c(hfrsNHCS, hfrsHJl, hfrsHNl), #hfrsHJn
                 Method=factor(rep(HFRtypes, each=length(hfrsNHCS)), levels=HFRtypes))

```

# Save everything
```{r}
us_dfs <- list(df_real_time, df_both, df_window_conv, 
               df_window_lag, df_lags, df_delays_9, df_delays_7,
               df_deaths_by_source, df_hfr_types, df_hfr_ests_by_source)
saveRDS(us_dfs, here::here(data_folder, "US_dfs.RData"))
```

