
```{r}
data_folder <- here::here("Data", "Real_data")
results_folder <- here::here("Data", "HFR_estimates")
# code_folder <- here::here("Code", "Analysis")
figure_folder <- here::here("Figures")
# source(here::here(code_folder, "helper_functions.R"))

# library(epidatr)
# library(stats)
library(patchwork)
library(ggplot2)
library(tidyverse)
```

```{r}
us_dfs <- readRDS(here::here(results_folder, "US_dfs.RData"))
df_real_time <- us_dfs[[1]]
df_both <- us_dfs[[2]]
df_window_conv <- us_dfs[[3]]
df_window_lag <- us_dfs[[4]]
df_lags <- us_dfs[[5]]
df_delays_9 <- us_dfs[[6]]
df_delays_7 <- us_dfs[[7]]
df_deaths_by_source <- us_dfs[[8]]
df_hfr_types <- us_dfs[[9]]
df_hfr_ests_by_source <- us_dfs[[10]]
```


```{r}
first_hosp_date <- as.Date("2020-07-28")
hosps <- readRDS(file.path(data_folder, "HHS_real_time.RData"))
hosps_final <- readRDS(file.path(data_folder, "HHS_finalized.RData"))
hosp_dates <- as.Date(names(hosps_final))
hfrs_nhcs <- readRDS(here::here(data_folder, "HFRs_NHCS_rescaled.RData"))
hfr_dates <- as.Date(names(hfrs_nhcs))
deaths <- readRDS(file.path(data_folder, "JHU_real_time.RData"))
death_dates <- as.Date(names(deaths))

d <- 75

est_weeks <- seq(first_hosp_date+d+14, min(max(hosp_dates)-4*7, max(hfr_dates)), by=7)
n_ests <- length(est_weeks)
```

### Properly-versioned data

```{r}
# Identify valid weeks. 
n_days_after <- 2 # Most recent date often not posted for over a week
week_st_idx <- 5 # HHS first issued Nov 16, 2020 (est_week[4]+1); 
est_weeks_realtime <- est_weeks[week_st_idx:n_ests] 
```


```{r}
## Plot just results with real-time counts
waves <- c("Original", "Delta", "Omicron") #Alpha vs Original
periods <- list(#est_weeks; as.Date("2021-05-01")
  as.Date(intersect(seq(as.Date("2020-11-01"), as.Date("2021-03-28"), by="day"), est_weeks_realtime)),
  as.Date(intersect(seq(as.Date("2021-06-01"), as.Date("2021-10-01"), by="day"), est_weeks_realtime)),
  as.Date(intersect(seq(as.Date("2022-01-01"), as.Date("2022-06-01"), by="day"), est_weeks_realtime))
)
# Convert to dataframe
shades <- do.call(rbind, lapply(periods, function(x) data.frame(xmin = x[1], xmax = x[length(x)], ymin = -Inf, ymax = Inf)))
```


```{r}
df_real_time = df_real_time |>
  mutate(Method = case_match(Method, "Approx. GT" ~ "Approximate GT", .default = Method))

first_xaxis_date <- as.Date("2021-01-01") #as.Date("2020-10-01")
rtPlot <- ggplot(data=df_real_time, aes(x = Date, y = HFR, color = Method, linetype=Method)) + 
  geom_rect(data = shades, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            inherit.aes = FALSE, fill = "lightgray", alpha = 0.5) +
  geom_line() + labs(color = "HFR", linetype = "HFR") + 
  ggtitle("Ratio estimates and approximate ground truth")+#, subtitle="Using real-time counts") +
  scale_x_date(breaks = seq(first_xaxis_date, max(df_real_time$Date), by = "3 months"), 
               date_labels = "%b %Y") +
  xlab("") + theme_bw() +
  theme(legend.position="bottom",
        plot.title=element_text(size=17),
        # plot.subtitle=element_text(size=13),
        legend.text=element_text(size=14),
        legend.title=element_blank())

rtPlot
ggsave(file.path(figure_folder, "Real", "US_ests_realtime.pdf"),
       plot = rtPlot, device = "pdf", width = 10, height = 5)
```

Plot results with both

```{r}
df_both = df_both |>
  mutate(type = case_match(type, 
                           "Ground truth" ~ "Approximate GT", 
                           "Convolutional" ~ "Convolutional ratio",
                           "Lagged" ~ "Lagged ratio"),
         source = case_match(source, 
                           "GT HFR" ~ "GT", 
                           "Finalized" ~ "Finalized",
                           "Real-Time" ~ "Real-time"))
         
bothPlot <- ggplot(data=df_both, aes(x = Date, y = HFR, color = type, linetype=source)) +
  geom_line(aes(group = interaction(source, type))) +
  ggtitle("Ratio estimates and approximate ground truth", subtitle="Using both real-time and finalized counts") +
  scale_x_date(breaks = seq(as.Date("2020-10-01"), max(df_both$Date), by = "3 months"),
               date_labels = "%b %Y") +
  xlab("") + theme_bw() +
  scale_linetype_manual(values=c("GT" = "solid", "Finalized" = "dashed", "Real-time" = "dotted"),
                        breaks=c("Finalized", "Real-time")) +
  theme(legend.position="bottom",
        plot.title=element_text(size=17),
        plot.subtitle=element_text(size=14),
        legend.text=element_text(size=14),
        legend.spacing.y = unit(0.5, "cm"),
        legend.margin = margin(t = -10),
        legend.box = "vertical",
        legend.title=element_blank())

bothPlot
ggsave(file.path(figure_folder, "Real", "US_ests_realtime_both.pdf"),
       plot = bothPlot, device = "pdf", width = 10, height = 5)
```

## Zoom in to explain bad windows

```{r}
hfrsHJn_rt <- df_real_time[df_real_time$Method=="Convolutional ratio", "HFR"]
hfrsHJl_rt <- df_real_time[df_real_time$Method=="Lagged ratio", "HFR"]
hfrsNHCS <- sapply(est_weeks, function(date) {
  mean(hfrs_nhcs[hfr_dates %in% seq(date - 6, date, by="day")], na.rm = TRUE)
}); names(hfrsNHCS) <- est_weeks
hfrsNHCS_rt <- hfrsNHCS[week_st_idx:n_ests]

plot_list <- list()
y1name <- ""; y2name <- ""#Hospitalizations
for (i in 1:length(waves)) {
  wave <- waves[i]
  bad_dates <- periods[[i]]
  conv_hfrs_period <- hfrsHJn_rt[est_weeks_realtime %in% bad_dates]
  lagged_hfrs_period <- hfrsHJl_rt[est_weeks_realtime %in% bad_dates]
  hosps_period <- hosps[hosp_dates %in% bad_dates]
  gt_hfrs_period <- hfrsNHCS_rt[est_weeks_realtime %in% bad_dates]
  deaths_period <- deaths[death_dates %in% bad_dates]

  dfAgg <- data.frame(Date=bad_dates,
                      Hosps = hosps_period,
                      Conv_HFRs=conv_hfrs_period,
                      Lagged_HFRs=lagged_hfrs_period,
                      GT_HFRs=gt_hfrs_period)
  dfAgg$Legend_Conv_HFR <- "Convolutional ratio"
  dfAgg$Legend_Lagged_HFR <- "Lagged ratio"
  dfAgg$Legend_HHS <- "Hospitalizations"
  dfAgg$Legend_GT_HFR <- "Approximate GT"
  
  ymin <- min(hfrsNHCS_rt, hfrsHJl_rt, hfrsHJn_rt)
  ymax <- max(hfrsNHCS_rt, hfrsHJl_rt, hfrsHJn_rt)
  scale <- (ymax - ymin) / (max(hosps) - min(hosps))
  a <- min(hosps)
  b <- ymin
  
 # scale <- min(lagged_hfrs_period) / mean(hosps_period) 
 # ymax <- max(lagged_hfrs_period, gt_hfrs_period, conv_hfrs_period, hosps_period*scale)*1.1
  
  p <- ggplot(data = dfAgg, aes(x = Date)) +
    # ggtitle(paste0("HFRs and Hospitalizations, ", wave)) +
    ggtitle(wave) +
    geom_line(data = dfAgg, aes(y = Conv_HFRs, color = Legend_Conv_HFR, linetype = Legend_Conv_HFR)) +
    geom_line(data = dfAgg, aes(y = Lagged_HFRs, color = Legend_Lagged_HFR, linetype = Legend_Lagged_HFR)) +
    geom_line(data = dfAgg, aes(y = GT_HFRs, color = Legend_GT_HFR, linetype = Legend_GT_HFR)) +
    geom_line(data = dfAgg, aes(y = (Hosps - a) * scale + b, color=Legend_HHS, linetype=Legend_HHS)) +
    ylab(y1name) + 
    scale_linetype_manual(breaks = c("Approximate GT", "Convolutional ratio", "Lagged ratio", "Hospitalizations"),
                        values = c("Approximate GT" = "solid", "Convolutional ratio" = "dotted", 
                                   "Lagged ratio"="dashed", "Hospitalizations" = "dotdash")) +
    scale_color_manual(breaks = c("Approximate GT", "Convolutional ratio", "Lagged ratio", "Hospitalizations"),
                     values = c("Approximate GT" = "red", "Convolutional ratio" = "green3", 
                                "Lagged ratio" = "dodgerblue", "Hospitalizations" = "black")) +
    scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +
    scale_y_continuous(#breaks = seq(from = ymin, to = ymax, by = 0.05), 
                       limits = c(ymin, ymax), 
                       sec.axis = sec_axis(~ (. - b) / scale + a, name = y2name)) +
    xlab("") +
    theme_bw() +
    theme(legend.position="bottom",
          axis.text.x = element_text(angle=45, vjust=0.5),
          plot.title = element_text(size=17),
          plot.margin = margin(0, 0, 0, 0, "pt"),
          legend.title=element_blank())
  plot(p)
  plot_list <- append(plot_list, list(p))
}
```

```{r}
wavePlot <- (plot_list[[1]] + plot_list[[2]] + plot_list[[3]]) +
  plot_layout(guides = "collect") &
  theme(legend.position = 'bottom',
        legend.text = element_text(size=14),
        legend.title=element_blank())

wavePlot
ggsave(file.path(figure_folder, "Real", "hfrs_by_wave.pdf"),
       plot = wavePlot, device = "pdf", width = 10, height = 4)
```

# Robustness checks

## Robustness to choice of hyperparameter

### Window size: JHU still messed up

Nishiura convolutional estimator

```{r, fig.width=10, fig.height=4}
df_window_conv = df_window_conv |>
  mutate(Window = recode_factor(Window, "Approx. GT" = "Approximate GT"))

p1 <- ggplot(df_window_conv, aes(x = Date, y = HFR, color = Window, linetype = Window)) +
  geom_line() + 
  xlab("") + ylab("HFR") +
  ggtitle("Varying window lengths, convolutional ratio") + theme_bw() +
  guides(color = guide_legend(nrow = 1)) +
  theme(legend.position="bottom",
        plot.title=element_text(size=17),
        legend.text=element_text(size=14),
        legend.title=element_blank())
p1

df_window_lag = df_window_lag |>
  mutate(Window = recode_factor(Window, "Approx. GT" = "Approximate GT"))

p2 <- ggplot(df_window_lag, aes(x = Date, y = HFR, color = Window, linetype = Window)) +
  geom_line() + 
  xlab("") + ylab("HFR") +
  ggtitle("Varying window lengths, lagged ratio") + theme_bw() +
  guides(color = guide_legend(nrow = 1)) +
  theme(legend.position="bottom",
        plot.title=element_text(size=17),
        legend.text=element_text(size=14),
        legend.title=element_blank())

ggsave(file.path(figure_folder, "Real", "window_size_conv.pdf"), 
       plot = p1, device = "pdf", width = 10, height = 4) 

ggsave(file.path(figure_folder, "Real", "window_size_lagg.pdf"), 
       plot = p2, device = "pdf", width = 10, height = 4) 
```

### Robustness to choice of lag

```{r}
df_lags = df_lags |> 
  mutate(Lag = recode_factor(Lag, "Approx. GT" = "Approximate GT"))

lagPlot <- ggplot(df_lags, aes(x = Date, y = HFR, color = Lag, linetype = Lag)) +
  geom_line() + theme_bw() +  
  xlab("") + ylab("HFR") + ggtitle("Varying lag parameters") +
  guides(color = guide_legend(nrow = 1)) +
  theme(legend.position="bottom",
        plot.title=element_text(size=17),
        legend.text=element_text(size=14),
        legend.title=element_blank())

lagPlot
ggsave(file.path(figure_folder, "Real", "hfrs_by_lag.pdf"),
       plot = lagPlot, device = "pdf", width = 10, height = 4)
```


## Robustness to choice of delay distribution

```{r}
frac <- 0.9
name <- "hfrs_by_delay1.pdf"
df_delays_9 = df_delays_9 |> 
  mutate(Method = recode_factor(Method, "GT" = "Approximate GT"))

DelayPlot <- ggplot(data=df_delays_9, aes(x = Date, y = HFR, color = Method, linetype = Method)) + 
  geom_line() + theme_bw() +  
  xlab("") + ylab("HFR") + ggtitle("Varying delay means, with larger variance") +
  guides(color = guide_legend(nrow = 1)) +
  theme(legend.position="bottom",
        plot.title=element_text(size=17),
        legend.text=element_text(size=14),
        legend.title=element_blank())

DelayPlot
ggsave(file.path(figure_folder, "Real", name), plot = DelayPlot, 
       device = "pdf", width = 10, height = 4)
```

```{r}
frac <- 0.7
name <- "hfrs_by_delay2.pdf"
df_delays_7 = df_delays_7 |> 
  mutate(Method = recode_factor(Method, "GT" = "Approximate GT"))

DelayPlot2 <- ggplot(data=df_delays_7, aes(x = Date, y = HFR, color = Method, linetype = Method)) + 
  geom_line() + theme_bw() +  
  xlab("") + ylab("HFR") + ggtitle("Varying delay means, with smaller variance") +
  guides(color = guide_legend(nrow = 1)) +
  theme(legend.position="bottom",
        plot.title=element_text(size=17),
        legend.text=element_text(size=14),
        legend.title=element_blank())

DelayPlot2
ggsave(file.path(figure_folder, "Real", name), plot = DelayPlot2, 
       device = "pdf", width = 10, height = 4) 
```

## Compare different versions of ground truth

### Load NCHS deaths and compute HFRs

## Display real-time counts

```{r}
deathSourcePlot <- ggplot(df_deaths_by_source, aes(x=Date, y=Deaths, color=Source, linetype=Source)) + 
  geom_line() + theme_bw() +
  ggtitle("Deaths by source, early 2022") + xlab("") +
  theme(legend.position="bottom", 
        legend.title=element_blank(),
        legend.text = element_text(size = 15),
        axis.title.y=element_text(size=14),
        plot.title = element_text(size = 17))
deathSourcePlot
ggsave(file.path(figure_folder, "Real", "death_curves.pdf"), 
       plot = deathSourcePlot, device = "pdf", width = 7, height = 5)
```


Compare ground truth

```{r}
df_hfr_types = df_hfr_types |>
  mutate(Method = recode_factor(Method, 
                                "NHCS Survey" = "NHCS survey",
                                "NCHS/HHS (retro)" = "NCHS/HHS",
                                "GT-Variants" = "Variant-based"))

altGT <- ggplot(data=df_hfr_types, aes(x = Date, y = HFR, color = Method, linetype = Method)) + 
  geom_line() + labs(color = "HFR", linetype = "HFR") + 
  ggtitle("Ground truth comparison") + 
  theme_bw() + xlab("") +
  theme(legend.position="bottom",
        plot.title=element_text(size=17),
        legend.text=element_text(size=14),
        legend.title=element_blank())

altGT
ggsave(file.path(figure_folder, "Real", "ApproxGT.pdf"), 
       plot = altGT, device = "pdf", width = 10, height = 4) 
```


Compare JHU and NCHS lagged ratios

```{r}
df_hfr_ests_by_source = df_hfr_ests_by_source |>
  mutate(Method = case_match(Method, 
                             "Approx. GT" ~ "Approximate GT",
                             "JHU Estimate" ~ "JHU-based",
                             "NCHS Estimate" ~ "NCHS-based"))

HFRsourcePlot <- ggplot(data=df_hfr_ests_by_source, aes(x = Date, y = HFR, color = Method, linetype=Method)) + 
  geom_line() + labs(color = "HFR", linetype = "HFR") + 
  ggtitle("Death data comparison") +
  scale_x_date(breaks = seq(as.Date("2020-10-01"), max(df_hfr_ests_by_source$Date), by = "3 months"), 
               date_labels = "%b %Y") +
  xlab("") + theme_bw() +
  theme(legend.position="bottom",
        plot.title=element_text(size=17),
        legend.text=element_text(size=14),
        legend.title=element_blank())

HFRsourcePlot
ggsave(file.path(figure_folder, "Real", "jhu_vs_nchs.pdf"), 
       plot = HFRsourcePlot, device = "pdf", width = 10, height = 4) 
```

